---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# hillshader

<!-- badges: start -->
[![R-CMD-check](https://github.com/pierreroudier/hillshader/workflows/R-CMD-check/badge.svg)](https://github.com/pierreroudier/hillshader/actions)
[![CRAN status](https://www.r-pkg.org/badges/version/hillshader)](https://CRAN.R-project.org/package=hillshader)
[![Total_Downloads](http://cranlogs.r-pkg.org/badges/grand-total/hillshader)](https://cran.r-project.org/package=hillshader)
<!-- badges: end -->

The `hillshader` package is a wrapper around the `rayshader` and `raster` packages to create hillshade relief maps using ray-tracing, and write them to (spatial) files.

## Installation

The stable version of `hillshader` can be installed from CRAN:

```r
install.packages("hillshader")
```

Alternatively, if you feel brave, you can install the development version of `hillshader` with the `remotes` package:

``` r
remotes::install_github("pierreroudier/hillshader")
```

## First steps

Below is a quick tutorial of the `hillshader` capabilities:

### The `hillshader` function

The `hillshader` function is the main function of that package, and allows to create a hillshade map as a `RasterLayer`:

```{r load_libs}
library(raster)
library(rayshader)

library(hillshader)
```

Note that the `hillshader` package includes the `maungawhau` and `maungawhau_hr` datasets. These are geo-referenced, raster datasets. `maungawhau` corresponds to the well-known `volcano` dataset. It is a $87 \times 61$ elevation  matrix for [Maungawhau](https://en.wikipedia.org/wiki/Maungawhau_/_Mount_Eden), one of the circa 80 volcanoes in the Auckland volcano filed, in Aotearoa/New Zealand. The `maungawhau_hr` dataset is a "high-resolution" version of that dataset, and is a 1 m resolution, $860 \times 600$ elevation matrix derived from a LiDAR dataset recorded by NZ Aerial Mapping & Aerial Surveying Limited for Auckland Council, and [distributed by Land Information New Zealand](https://data.linz.govt.nz/layer/53405-auckland-lidar-1m-dem-2013/). 

```{r mwh}
layout(matrix(c(1,2), nrow = 1, ncol = 2))
image(maungawhau, asp = 1, main = "Maungawhau", col = terrain.colors(100))
image(maungawhau_hr, asp = 1, main = "Maungawhau (high-resolution)", col = terrain.colors(100))
```

The `hillshader` function can be simply called on a elevation raster to generate a hillshade `RasterLayer`. By default, the shader used is `rayshader::ray_shade`, with its default values.

```{r hillshader}
hs <- hillshader(maungawhau_hr)
plot_map(hs)
```

### Add more shaders!

The `hillshader` function accept a `shader` option, with is a list of the successive shader functions to apply to create the hillshade layer. The accepted values must be `rayshader` shader functions (`ray_shade`, `ambient_shade`, `lamb_shade`), and the order is important.

```{r hillshader2}
hs <- hillshader(
  elevation = maungawhau_hr, 
  shader = c("ray_shade", "ambient_shade")
)

plot_map(hs)
```

### Changing sun position

The `hillshader` function uses the `rayshader` options defaults, but other values can be specify and passed as arguments:

```{r hillshader3}
hs <- hillshader(
  elevation = maungawhau_hr, 
  shader = c("ray_shade", "ambient_shade"),
  sunangle = 180,
  sunaltitude = 25
)

plot_map(hs)
```

### How does it compare to your usual hillshade method?

```{r hs-compare}
library(raster)

slope <- terrain(maungawhau_hr, out = "slope")
aspect <- terrain(maungawhau_hr, out = "aspect")
hs_raster <- hillShade(
  slope, 
  aspect, 
  angle = 40, 
  direction = 325
)

hs_hillshader <- hillshader(
  maungawhau_hr, 
  c("ray_shade", "ambient_shade"),
  sunangle = 325,
  sunaltitude = 40
)

layout(matrix(c(1,2), nrow = 1, ncol = 2))
image(hs_raster, asp = 1, main = "Classic hillshade", col = grey.colors(100))
image(hs_hillshader, asp = 1, main = "Ray-traced hillshade", col = grey.colors(100))
```

### Saving to file

If a `filename` is passed to `hillshader`, then the resulting hillshade layer is saved to file. This is a wrapper around `raster::writeRaster`, and options specific to the latter function can be used.

```{r hillshader4, eval=FALSE}
hillshader(
  elevation = maungawhau_hr, 
  shader = c("ray_shade", "ambient_shade"),
  sunangle = 180,
  sunaltitude = 25,
  filename = "hillshade.tif"
)
```

## Advanced use in the `rayshader` pipelines

The `hillshader` package provides three functions that can be used within the `rayshader` pipelines:

- `add_shadow_2d`: a function that multiplies a shadow map by another shadow map, a corrected 2D version of `rayshader::add_shadow`,
- `matrix_to_raster`: a function that converts a matrix (typically used by the `rayshader` functions) back to a `RasterLayer`, for input into a GIS workflow,
- `write_raster`: a function that a hillshade matrix to a raster file format.

```{r advanced, eval=FALSE}
library(rayshader)
library(hillshader)

# Create elevation matrix
el_mat <- raster_to_matrix(maungawhau_hr)

el_mat %>%
  # Create hillshade layer using 
  # ray-tracing
  ray_shade %>%
  # Add ambient shading
  add_shadow_2d(
    ambient_shade(
      heightmap = el_mat
    )
  ) %>% 
  # Write to GIS file
  write_raster(
    elevation = maungawhau_hr,
    filename = "hillshade.tif"
  )
```
